# EP17 - Detection Capabilities
Evaluation of Detection Capability for Clinical Laboratory Measurement Procedures

This work is based on EP17-A2, June 2012.

Simple, elegant code with minimal duplication that can be adapted for other uses.

## Appendix A
Evaluation of Limits of Blank and Detection by the Classical Approach

```{r Setup, echo = FALSE}
suppressPackageStartupMessages(
  library(tidyverse)
)
```

### Table A1
Observed Blank Sample Results for Reagent Lot 1 (Units are pg/mL)
```{r Table A1}
# Table A1. Observed Blank Sample Results for Reagent Lot 1 (Units are pg/mL)
table_a1 <- tibble(
  day = rep(1:3, each = 4),
  replicate = rep(1:4, times = 3),
  `Blank 1` = c( 2.6, -0.8, 5.5, 6.0, 4.5, 0.6, -2.3,   3.4, 5.9, 7.6,  4.1, -1.4),
  `Blank 2` = c( 1.0,  2.9, 4.9, 8.0, 6.9, 5.0,  3.4,   1.2, 6.5, 5.6, -2.2,  2.3),
  `Blank 3` = c(-4.4, -3.4, 7.0, 6.9, 4.3, 3.2, -1.4,  -4.2, 5.9, 7.6,  3.8,  5.8),
  `Blank 4` = c( 1.5, -1.9, 5.1, 5.7, 4.1, 4.5, -0.6,   0.5, 5.4, 7.6,  4.4,  6.6),
  `Blank 5` = c( 1.2, -0.7, 6.1, 5.1, 4.8, 3.3, -2.8,  -1.4, 8.7, 3.6,  5.1,  3.5)
)

```

### Table A2
Observed Low Level Sample Results for Reagent Lot 1 (Units are pg/mL)
```{r Table A2}
# Table A2. Observed Low Level Sample Results for Reagent Lot 1 (Units are pg/mL)
table_a2 <- tibble(
  day = rep(1:3, each = 4),
  replicate = rep(1:4, times = 3),
  `Low 1` = c(21.0, 22.8, 28.2, 25.9, 26.4, 28.3, 20.7, 21.9, 24.7, 22.5, 28.5, 29.2),
  `Low 2` = c(13.3, 12.6, 18.2, 14.7, 17.8, 14.0, 14.1, 12.5, 11.3, 12.2, 16.2, 13.9),
  `Low 3` = c(12.8, 12.9, 17.4, 16.0, 15.9, 14.1, 11.3,  9.4, 10.6, 13.6, 17.6, 14.9),
  `Low 4` = c(17.3, 19.2, 21.5, 22.2, 24.1, 25.8, 16.0, 16.4, 24.9, 23.8, 22.1, 26.1),
  `Low 5` = c(19.2, 22.7, 28.3, 26.2, 25.1, 30.3, 23.4, 19.2, 26.3, 23.1, 27.5, 30.1)
)
```

### Table A3
Observed Blank Sample Results for Reagent Lot 2 (Units are pg/mL)
```{r Table A3}
# Table A3. Observed Blank Sample Results for Reagent Lot 2 (Units are pg/mL)
table_a3 <- tibble(
  day = rep(1:3, each = 4),
  replicate = rep(1:4, times = 3),
  `Blank 1` = c(4.6,  4.1,  1.6, 3.7, 2.2, 0.7, 4.6, 2.6, 1.1, -4.4, 0.9,   0.7),
  `Blank 2` = c(9.2,  8.3,  4.8, 5.4, 4.8, 6.3, 5.4, 9.6, 7.7,  3.1, 6.1,  10.0),
  `Blank 3` = c(6.1,  3.2,  3.9, 1.4, 3.1, 4.1, 1.0, 3.4, 0.1,  0.4, 2.9,  -1.6),
  `Blank 4` = c(4.0, 11.5,  4.5, 3.6, 4.4, 6.8, 7.1, 4.2, 3.7,  3.7, 5.3,   4.5),
  `Blank 5` = c(4.0,  6.2, -0.2, 2.3, 1.6, 2.6, 6.4, 5.7, 4.2,  3.7, 1.4,   1.5) 
)
```

### Table A4
Observed Low Level Sample Results for Reagent Lot 2 (Units are pg/mL)
```{r Table A4}
# Table A4. Observed Low Level Sample Results for Reagent Lot 2 (Units are pg/mL)
table_a4 <- tibble(
  day = rep(1:3, each = 4),
  replicate = rep(1:4, times = 3),
  `Low 1` = c(22.0, 22.5, 21.8, 22.1, 20.3, 21.0, 25.3, 26.0, 27.2, 25.1, 25.3, 25.3),
  `Low 2` = c(15.6, 21.2, 14.8, 14.9, 16.0, 15.8, 21.6, 22.8, 15.3, 18.7, 18.3, 19.5),
  `Low 3` = c(13.0, 15.9,  9.0,  7.0, 13.4,  8.5, 16.3, 18.1, 12.4, 11.1, 11.3, 10.1),
  `Low 4` = c(18.8, 17.6, 14.1, 14.9, 19.2, 15.8, 19.8, 21.4, 18.0, 18.0, 19.6, 23.1),
  `Low 5` = c(32.9, 30.4, 29.4, 27.6, 27.7, 30.6, 31.4, 30.4, 32.5, 28.9, 29.8, 35.1)
)
```

### Tidy Table Cleanup
Before we move on, we can make our job significantly easier going forward by reorganizing our tables into a tidy, rowwise format. 

```{r Cleanup Tables A1-A4 into tidy format}

# Add the reagent lots, pivot to a longer format, and convert non-value columns to factors
tables_a14 <- map2_dfr(.x = list(table_a1, table_a2, table_a3, table_a4),  # The tables we need to clean up
                     .y = list("Lot 1", "Lot 1", "Lot 2", "Lot 2"),  # The lots of reagent that map to each of the tables
                     .f = ~ mutate(.x, reagent_lot = .y) %>%
                       pivot_longer(cols = starts_with(c("Blank", "Low")), 
                                    names_to = "sample", 
                                    values_to = "value")) %>%
  separate(sample, into = c("sample_type", "sample_number"), sep = " ", remove = FALSE) %>%
  mutate_at(vars(-value), as.factor)

glimpse(tables_a14)

```

### Table A5
Rank Positions and LoB from Blank Sample Test Results
```{r Table A5}

alpha <- 0.05  # Type I error risk
pct_B <- 1 - alpha

B <- nlevels(tables_a14$day) *        # Number of days
  nlevels(tables_a14$sample_number) * # Number of unique samples
  nlevels(tables_a14$replicate)       # Number of replicates

rank_position <- 0.5 + (B * pct_B)

# Reproduce table A5  
tables_a14 %>%
  group_by(reagent_lot, sample_type) %>%
  mutate(rank_position = rank(value, ties.method = "random")) %>%
  filter(sample_type == "Blank",
        rank_position %in% 56:60) %>% 
  ungroup() %>%
  arrange(sample_type, reagent_lot, rank_position) %>%
  select(rank_position, reagent_lot, value) %>%
  pivot_wider(names_from = reagent_lot, values_from = value)

# Type 2 quantiles provide averaging at discontinuities
limits_nonparametric <- tables_a14 %>%
  group_by(reagent_lot, sample_type) %>%
    filter(sample_type == "Blank") %>%
  summarize(LoB = quantile(value, 1 - alpha, type = 2), .groups = "drop")  

limits_nonparametric
```


### Table A6
SDs and LoD Calcuations From Low Level Sample Test Results
```{r Table A6}

z <- qnorm(1-alpha)  # 95th percentile of normal distribution
#z <- round(z, digits = 3)

# Degrees of freedom, total measurements less the number of unique samples
df <- B - nlevels(tables_a14$sample_number)

# "Multiplier factor" as per eqn. 6
c_p = z / (1 - ( 1 / (4 * df)))

table_a6 <- tables_a14 %>%
  filter(sample_type == "Low") %>%
  group_by(sample, reagent_lot) %>%
  summarize(n = n(), SD = sd(value), .groups = "drop") %>%
  arrange(reagent_lot)

table_a6

limits_nonparametric <- table_a6 %>% 
  group_by(reagent_lot) %>%
  summarize(SD_L = sqrt(sum((n - 1) * SD^2) / sum(n-1)), .groups = "drop") %>%
  left_join(limits_nonparametric, by = "reagent_lot") %>%
  mutate(LoD = round(LoB + c_p * SD_L, 1))

## I believe there to be an error in this table - CD, 16-July-2020

limits_nonparametric

```

### Figure A1
Histogram of Combined Blank Sample Measurements
```{r Figure A1}

tables_a14 %>%
  filter(sample_type == "Blank") %>%
  ggplot(aes(x = value)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "lightgrey",
                 boundary = 0.5, closed = "left") +  # Match the binning-style that CLSI uses
  scale_x_continuous(breaks = seq(from = -6, to = 12, by = 2), limits = c(-6, 12), name = "") +
  scale_y_continuous(breaks = seq(2, 10, by = 2), name = "Count")

shapiro.test(tables_a14 %>%
  filter(sample_type == "Blank") %>%
  pull(value) 
)
```

### Table A7
LoB Calculations Using Parametric Data Analysis Options
```{r Table A7}

table_a7 <- tables_a14 %>%
  filter(sample_type == "Blank") %>%
  group_by(reagent_lot) %>%
  summarize(M_B = round(mean(value), 2),
            SD_B = round(sd(value), 2),
            c_p = c_p,
            .groups = "drop") %>%
  mutate(LoB = round(M_B + c_p * SD_B, 1))

# Row 1 of this table doesn't allign with the document
table_a7

```


```{r Calculate LoD parametrically}
limits_parametric <- table_a6 %>% 
  group_by(reagent_lot) %>%
  summarize(SD_L = sqrt(sum((n - 1) * SD^2) / sum(n-1)), .groups = "drop") %>%
  right_join(table_a7, by = "reagent_lot") %>%
  mutate(LoD = round(LoB + c_p * SD_L, 1))

limits_parametric


```


```{r Appendix A Error Checking, echo = FALSE, include = FALSE}
# Confirm we've reproduced the values in the text
stopifnot(
  near(
    limits_nonparametric$LoB, 
    c(7.6, 9.4)
  ))

stopifnot(
  near(
    limits_nonparametric$LoD,
    c(12.7, 13.8)    # Values according to my math  
    # c(14.5, 13.8)  # Values according to document
  ))

stopifnot(
  near(
    limits_parametric$LoB,
    c(8.8, 8.6)
  ))

stopifnot(
  near(
    limits_parametric$LoD,
    c(13.9, 13.0)    # Values according to my math  
    # c(13.9, 13.2)  # Values according to document
  ))
```

## Appendix B
Evaluation of Limit of Detection by the Precision Profile Approach

### Table B1
Observed Within-Laboratory Precision and Mean Concentration for All Samples

```{r Table B1}

table_b1 <- tibble(
  sample_id = LETTERS[1:6],
  mean.lot1 = c(0.69, 1.42, 2.65, 4.08, 6.08, 10.36),
  SD_WL.lot1 = c(0.39, 0.39, 0.46, 0.55, 0.64, 1.12),
  mean.lot2 = c(0.78, 1.73, 2.89, 3.82, 6.33, 10.92), 
  SD_WL.lot2 = c(0.29, 0.54, 0.55, 0.63, 0.82, 1.38))

table_b1

```

### Tidy Table Cleanup
```{r Tidy Table Cleanup, Table B1, rows.print = 12}

table_b1 <- table_b1 %>%
  pivot_longer(-sample_id) %>%
  separate(name, sep = "\\.", into = c("name", "reagent_lot")) %>%
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(reagent_lot = str_replace(reagent_lot, "lot", "Lot ")) %>%
  mutate_if(is.character, factor)

glimpse(table_b1)
```

### Figure B1
Precision Profiles of Within-Laboratory Precision (SD_WL) vs Measurand Concentration With Second-Order Polynomial Model Fit Overlays

```{r Figure B1}

fits <- table_b1 %>% 
  group_by(reagent_lot) %>%
  do(poly_fit = lm(formula = .$SD_WL ~ poly(.$mean, 2, raw = TRUE)))

# These are the same coefficients that Excel returns
# Presumably this is what the guideline means when it says 'done using a spreadsheet'
# They differ from the coefficients reported in the guideline

fits$poly_fit

# Lot 1 Reported:    0.3741 / 0.0149 / 0.0055
# Lot 1 Calculated:  0.3756 / 0.0137 / 0.0056
# Lot 2 Reported:    0.2801 / 0.0817 / 0.0017
# Lot 2 Calculated:  0.3086 / 0.0747 / 0.0020

table_b1 %>%
  ggplot(aes(x = mean, y = SD_WL)) +
  geom_point(aes(color = reagent_lot, shape = reagent_lot), size = 3) +
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), aes(color = reagent_lot, lty = reagent_lot), se = FALSE) +
  scale_shape_manual(values = c("diamond", "square"), name = "") +
  scale_color_discrete(name = "") +
  scale_x_continuous(name = "[Measurand] (ng/mL)",   limits = c(0, 12),  breaks = seq(0, 12,  by = 2)) +
  scale_y_continuous(name = "Within-Lab SD (ng/mL)", limits = c(0, 1.6), breaks = seq(0, 1.6, by = 0.2)) +
  scale_linetype_manual(values = c(1, 2), name = "")

```

### Equations B1-B3
```{r Equations B1-B3}

# Unable to calculate directly (either in R or via Excel), so input from text
coefficients <- tribble(
  ~ reagent_lot, ~ C0, ~ C1, ~ C2,
  "Lot 1", 0.3741, 0.0149, 0.0055, # From text
# "Lot 1", 0.3756, 0.0137, 0.0056, # Calculated
  "Lot 2", 0.2801, 0.0817, 0.0017) # From text
# "Lot 2", 0.3086, 0.0747, 0.0020) # Calculated


N <- 80  # Measurements per sample
K <- 6   # Number of samples
N_tot <- N * K  # Total measurements
alpha <- 0.05
Z <- qnorm(1 - alpha)
c_p <- Z / ( 1 - ( 1 / (4 * (N_tot - K))))

LoB <- 0.51  # From section 5.4.3.2

LoD <- coefficients %>%
  mutate(LoD = LoB + c_p * (C0 + C1 * LoB + C2 * LoB^2))

```

### Table B2
Trial LoD Values vs MC
```{r Table B2, rows.print = 19}

table_b2 <- tibble(
  mc = c(seq(from = 0.5, to = 1.3, by = 0.1), 
         seq(from = 1.1, to = 1.2, by = 0.01)),
  SD_WL.lot1 = c(0.383, 0.385, 0.387, 0.390, 0.392, 0.395, 0.397, 0.400, 0.403, 
                 0.397, 0.397, 0.398, 0.398, 0.398, 0.399, 0.399, 0.399, 0.399, 0.400, 0.400),
  SD_WL.lot2 = c(0.348, 0.355, 0.363, 0.371, 0.379, 0.387, 0.394, 0.402, 0.410,
                 0.394, 0.395, 0.396, 0.397, 0.398, 0.398, 0.399, 0.400, 0.401, 0.401, 0.402))

# Tidy the table
table_b2 <- table_b2 %>%
  distinct(mc, .keep_all = TRUE) %>%  # Remove the duplicate rows
  pivot_longer(-mc) %>%
  separate(name, into = c("name", "reagent_lot"), sep = "\\.") %>%
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(reagent_lot = str_replace(reagent_lot, "lot", "Lot ")) %>%
  mutate_if(is.character, factor) %>%
  arrange(reagent_lot)

# Calculate the trial LoD and the bias at each concentration
table_b2 <- table_b2 %>%
  mutate(trial_LoD = round(LoB + c_p * SD_WL, 2),
         bias = round(trial_LoD - mc, 2)) 

table_b2

# LoDs are where there is no bias
table_b2 %>%
  filter(!bias) %>%
  select(reagent_lot, LoD = trial_LoD)

```

## Appendix C
Evaluation of Limit of Detection by the Probit Approach

### Table C1
Observed Proportions of Positive Test Results With the Planned Dilutions
```{r Table C1}

table_c1 <- tibble(
  concentration = c(0.000, 0.025, 0.050, 0.150, 0.300, 0.500),
  positive_total.lot1 = c("0/22", "23/32", "29/32", "32/32", "32/32", "32/32"),
  positive_total.lot2 = c("0/22", "28/32", "32/32", "32/32", "32/32", "32/32"),
  positive_total.lot3 = c("0/22", "27/32", "32/32", "32/32", "32/32", "32/32"),
  hit_rate.lot1 = map_dbl(positive_total.lot1, ~ eval(parse(text = .))),  # This is a quick (lazy) way of
  hit_rate.lot2 = map_dbl(positive_total.lot2, ~ eval(parse(text = .))),  # evaluating the fractions in
  hit_rate.lot3 = map_dbl(positive_total.lot3, ~ eval(parse(text = .))))  # the first three columns.

table_c1
```

### Table C2
Observed Proportions of Positive Test Results With Additional Dilutions
```{r Table C2}

table_c2 <- tibble(
  concentration = c(0.006, 0.014),
  positive_total.lot1 = c("11/30", "15/30"),
  positive_total.lot2 = c("12/30", "22/30"),
  positive_total.lot3 = c("22/34", "31/34"),
  hit_rate.lot1 = map_dbl(positive_total.lot1, ~ eval(parse(text = .))),  # This is a quick (lazy) way of
  hit_rate.lot2 = map_dbl(positive_total.lot2, ~ eval(parse(text = .))),  # evaluating the fractions in
  hit_rate.lot3 = map_dbl(positive_total.lot3, ~ eval(parse(text = .))))  # the first three columns.

table_c2
```

### Tidy Table Cleanup
```{r Table C2}

tables_c12 <- table_c1 %>%
  bind_rows(table_c2) %>% 
  select(-starts_with("positive_total.lot")) %>%
  pivot_longer(-concentration) %>%
  separate(name, into = c("name", "reagent_lot"), sep = "\\.") %>%
  pivot_wider(names_from = name, values_from = value) %>%
  mutate(reagent_lot = str_replace(reagent_lot, "lot", "Lot ")) %>%
  arrange(reagent_lot, concentration)
```


### Table C3
```{r Table C3}

df <- tables_c12 %>%
  filter(reagent_lot == "Lot 2")

probit_model <- glm(formula = df$hit_rate ~ df$concentration, family = quasibinomial(link = "probit"))

intercept <- coef(probit_model)[[1]]
slope <- coef(probit_model)[[2]]

predict(probit_model, newdata = tibble("concentration" = 0.05), type = "response")

invlink_fn <- function(x) pnorm(x * slope + intercept)

df %>%
  ggplot(aes(x = concentration, y = hit_rate)) +
  geom_point() +
  scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1), limits = c(0.001, 10)) +
  geom_smooth(formula = y ~ x, color = "black", fullrange = TRUE,
              method = "glm", method.args = list(family = quasibinomial(link = "probit"))) +
  stat_function(fun = invlink_fn, color = "red") +
  geom_hline(yintercept = 0.95, lty = 2) +
  geom_vline(xintercept = 0.033, lty = 2) +
  coord_cartesian(xlim = c(0.001, 0.5))


p <- ggplot(df, aes(x = concentration, y = hit_rate)) + 
    geom_smooth(formula = y ~ x, color = "black", fullrange = TRUE,
              method = "glm", method.args = list(family = quasibinomial(link = "probit")))



```
