```{r Setup, echo = FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
})

options(dplyr.summarise.inform = FALSE)
```

# EP05 - Precision
Evaluation of Precision of Quantitative Measurement Procedures

This work is based on EP05-A3, October 2014

Simple, elegant code with minimal duplication that can be adapted for other uses.

## Appendix A
Single-Site Study


### Table A1
Glucose Precision Evaluation Measurements (mg/dL)
```{r Table A1}

# Table A1. Glucose Precision Evaluation Measurements (mg/dL)
table_a1 <- tibble(
  day = 1:20,
  run1.rep1 = c(242, 243, 247, 249, 246, 244, 241, 245, 243, 244, 252, 249, 242, 246, 247, 240, 241, 244, 241, 247),
  run1.rep2 = c(246, 242, 239, 241, 242, 245, 246, 245, 239, 246, 251, 248, 240, 249, 248, 238, 244, 244, 239, 240),
  run2.rep1 = c(245, 238, 241, 250, 243, 251, 245, 243, 244, 247, 247, 251, 251, 248, 245, 239, 245, 237, 247, 245),
  run2.rep2 = c(246, 238, 240, 245, 240, 247, 247, 245, 245, 239, 241, 246, 245, 240, 246, 242, 248, 242, 245, 242))

```

#### Normalization
```{r Table A1 Normalization}

table_a1 <- table_a1 %>%
  pivot_longer(cols = starts_with("run"), names_to = "run_rep", values_to = "result") %>%
  separate(run_rep, sep = "\\.", into = c("run", "rep")) %>%
  mutate_at(vars("run", "rep"), ~ str_remove(., "(run)|(rep)")) %>%
  mutate_at(vars(-result), factor)

glimpse(table_a1)

```

### Figure A1
Glucose Precision Evaluation Results
```{r Figure A1}

table_a1 %>%
  ggplot(aes(x = day, y = result)) +
  geom_point(aes(color = run, shape = run)) +
  scale_x_discrete(name = "Test Day", breaks = seq(0, 20, by = 4)) +
  scale_y_continuous(name = "Glucose, mg/dL", breaks = seq(236, 254, by = 2)) +
  scale_color_discrete(name = "", labels = c("Morning Run", "Afternoon Run")) +
  scale_shape_manual(name = "", values = c("square", "circle"), labels = c("Morning Run", "Afternoon Run"))

```


### Table A2
ANOVA Summary Table for the Glucose Example

```{r Table A2}
table_a2 <- table_a1 %>%
  aov(result ~ day + day:run, data = .) %>%
  broom::tidy() %>%
  select(source = term, SS = sumsq, DF = df, MS = meansq) %>%
  column_to_rownames(var = "source")

table_a2
```

### Equations A1-A9

```{r Equations A1-A9}
x_bar <- mean(table_a1$result)

MS_error <- table_a2["Residuals", "MS"]
MS_run <- table_a2["day:run", "MS"]
MS_day <- table_a2["day", "MS"]

DF_error <- table_a2["Residuals", "DF"]
DF_run <- table_a2["day:run", "DF"]
DF_day <- table_a2["day", "DF"]

N <- nrow(table_a1)
n_rep <- nlevels(table_a1$rep)
n_run <- nlevels(table_a1$run)
n_day <- nlevels(table_a1$day)

# If any of these terms are less than zero, carry them forward as zero
V_error <- MS_error  # A1
V_run <- max((MS_run - MS_error)/n_rep, 0)  # A2
V_day <- max((MS_day - MS_run)/(n_run*n_rep), 0)  # A3

S_R <- sqrt(V_error)  # A4
S_WL <- sqrt(V_day + V_run + V_error)  # A5

# Only valid for a 20 x 2 x 2 study
alpha_error <- 0.5
alpha_run <- 0.25
alpha_day <- 0.25

df_R <- N - n_day*n_run  # A6

# A7
df_WL <- round(  
  ((alpha_day * MS_day) + (alpha_run * MS_run) + (alpha_error * MS_error))^2 
  / 
  ((alpha_day * MS_day)^2/DF_day + (alpha_run * MS_run)^2/DF_run + (alpha_error * MS_error)^2/DF_error)
)

CV_R <- S_R / x_bar * 100  # A8
CV_WL = S_WL / x_bar * 100  # A9

tibble(
  `Sample Description` = "Patient Pool",
  `Mean (mg/dL)` = x_bar,
  `Repeatability` = sprintf("%.2f (%.1f%%)", S_R, CV_R),
  `Within-Laboratory Precision` = sprintf("%.2f (%.1f%%)", S_WL, CV_WL)
) 
```

### Table A3
Inputs for calculating 95% CIs for the Glucose Example

```{r Table A3}
alpha <- 0.05

table_a3 <- tibble(
  value = c("Repeatability", "Within-Laboratory Reproducibility"),
  S = c(S_R, S_WL),
  DF = c(df_R, df_WL),
  chisq_lower = qchisq(1-alpha/2, DF),
  chisq_upper = qchisq(alpha/2, DF))

table_a3 
```

### Equations A10-A11
```{r Equations A10-A11}
table_a3 %>%
  mutate(S_lower = S * sqrt(DF/chisq_lower),  # A10 and A11
         S_upper = S * sqrt(DF/chisq_upper),
         CV_lower = S_lower / x_bar * 100,
         CV_upper = S_upper / x_bar * 100,
         CV_range = sprintf("%.1f%% - %.1f%%", CV_lower, CV_upper)) %>%
  select(value | starts_with("S") | starts_with("CV"))

# There is a rounding error in this table, the text reports 0.95% but without rounding, the calculated value is 0.944%
```

## Appendix B
CA19-9 Precision Evaluation Measurements (kU/L). Six samples, P1 to Q6, where P signifies patient sample pool and "Q" signifies QC material.

### Table B1
```{r Table B1}

# I'm not mad about having to type all these dumb numbers in. Not mad at all
table_b1 <- tibble(
  day = rep(rep(1:5, each = 5), times = 2*3),
  sample = rep(c("P1", "P2", "Q3", "Q4", "P5", "Q6"), each = 25),
  `Site 1` = c(
    12.5,  11.8,  11.6,  11.6,  11.7,  12.2,  11.8,  11.6,  12.2,  11.5,  12.0,  11.9,  11.4,  
    11.6,  11.9,  11.9,  11.1,  12.4,  11.3,  11.5,  13.3,  10.7,  10.3,  11.7,  11.9,   # P1
    45.0,  45.1,  42.4,  42.4,  42.5,  42.4,  41.7,  43.0,  41.8,  41.6,  45.1,  42.7,  41.6,  
    41.0,  42.2,  41.8,  44.4,  42.8,  41.4,  41.8,  41.2,  40.9,  41.1,  40.2,  40.9,   # P2
    59.0,  57.7,  58.1,  56.3,  55.6,  56.5,  56.3,  58.4,  57.3,  56.5,  57.7,  56.7,  58.6,  
    56.1,  55.2,  56.5,  58.4,  55.4,  55.7,  57.7,  56.9,  56.3,  56.3,  55.7,  55.0,   # Q3
    175.0, 167.3, 169.7, 167.8, 170.5, 167.3, 166.7, 172.0, 170.1, 170.5, 167.4, 168.8, 170.5, 
    176.5, 168.2, 173.3, 171.4, 164.1, 166.9, 167.4, 164.8, 167.1, 167.3, 165.8, 165.8,  # Q4
    386.6, 393.9, 392.8, 383.3, 388.0, 389.0, 394.3, 374.2, 379.8, 368.1, 381.0, 380.1, 375.3, 
    375.0, 375.0, 400.2, 386.2, 364.6, 371.1, 371.3, 385.1, 390.6, 386.2, 377.4, 384.9,  # P5
    433.9, 434.6, 424.9, 425.2, 424.9, 420.6, 437.5, 413.6, 416.2, 410.7, 423.0, 412.7, 426.7, 
    435.3, 417.9, 435.9, 424.9, 406.6, 425.8, 427.2, 418.0, 415.8, 409.2, 411.1, 423.2), # Q6
  `Site 2` = c(
    12.8,  12.8,  12.7,  13.2,  11.1,  11.4,  10.7,  11.0,  13.8,  11.4,  14.7,  13.7,  13.2, 
    13.2,  13.1,  13.6,  13.5,  12.8,  12.3,  12.0,  14.1,  12.8,  12.2,  15.8,  13.3,   # P1
    40.6,  39.3,  41.4,  40.3,  37.3,  38.2,  37.5,  38.7,  43.0,  40.7,  42.1,  41.3,  40.5,  
    38.8,  39.9,  40.9,  40.7,  39.6,  40.1,  39.5,  40.8,  40.6,  41.4,  41.4,  39.7,   # P2
    53.9,  55.1,  54.9,  53.7,  51.3,  52.7,  54.4,  49.8,  51.4,  49.0,  55.1,  55.0,  54.1,  
    52.5,  54.5,  55.2,  56.1,  55.2,  55.8,  54.9,  54.5,  55.9,  51.7,  50.0,  54.2,   # Q3
    156.6, 157.7, 155.8, 160.8, 153.6, 162.7, 157.1, 156.1, 149.5, 157.0, 159.9, 163.8, 160.9, 
    164.4, 156.3, 162.2, 164.6, 161.5, 159.2, 157.9, 158.2, 162.9, 162.4, 159.6, 160.0,  # Q4
    380.0, 382.5, 368.5, 379.7, 361.7, 366.8, 371.7, 363.3, 379.1, 366.7, 375.1, 382.1, 371.3, 
    374.3, 370.1, 372.7, 376.1, 363.9, 371.4, 370.8, 365.7, 371.2, 387.5, 385.5, 367.3,  # P5
    410.9, 416.3, 391.8, 390.7, 388.4, 401.2, 409.0, 397.0, 395.7, 384.1, 401.4, 411.9, 292.2, 
    409.3, 391.9, 395.9, 415.7, 400.5, 409.4, 390.8, 410.5, 399.7, 387.9, 398.9, 382.5), # Q6
  `Site 3` = c(
    12.2,  11.7,  11.6,  11.9,  11.6,  12.0,  11.7,  11.9,  12.1,  11.7,  12.0,  11.7,  11.5,  
    11.8,  11.7,  11.5,  11.8,  11.9,  12.3,  12.6,  12.0,  10.5,  10.5,  11.2,  11.1,   # P1
    45.1,  43.8,  42.4,  42.4,  42.1,  42.1,  42.4,  42.4,  43.5,  41.9,  43.9,  42.2,  41.3,  
    41.4,  42.0,  43.1,  43.6,  42.1,  41.3,  42.8,  41.1,  41.0,  40.7,  41.4,  43.9,   # P2
    58.4,  57.9,  57.2,  56.4,  56.1,  56.4,  57.4,  57.9,  57.5,  55.9,  57.2,  57.7,  57.4,  
    56.3,  56.5,  57.5,  56.9,  55.6,  56.3,  56.1,  56.6,  56.3,  56.0,  55.7,  57.0,   # Q3
    171.2, 168.5, 168.8, 167.6, 170.5, 167.0, 169.4, 171.1, 168.8, 169.4, 168.1, 169.7, 173.5,
    175.9, 167.8, 172.4, 167.8, 165.5, 165.9, 166.0, 166.0, 167.2, 166.6, 168.2, 168.8,  # Q4
    390.3, 393.4, 388.1, 386.2, 378.1, 391.7, 384.3, 377.0, 380.4, 371.6, 380.6, 377.7, 375.2, 
    387.6, 373.2, 393.2, 375.4, 367.9, 378.1, 378.3, 387.9, 388.4, 381.8, 382.7, 383.7,  # P5
    434.3, 429.8, 425.1, 422.9, 417.8, 429.1, 425.6, 414.9, 419.6, 414.3, 417.9, 419.7, 431.0, 
    435.6, 422.6, 430.4, 415.8, 416.2, 421.9, 415.2, 416.9, 412.5, 410.2, 418.0, 415.2), # Q6
)

```

#### Normalization

```{r Table B1 Normalization}
table_b1 <- table_b1 %>%
  pivot_longer(cols = starts_with("Site"), names_to = "site", values_to = "result") %>%
  mutate(site = str_remove(site, "Site ")) %>%
  mutate_at(vars(-result), as.factor)
```
