---
output:
  pdf_document: default
  html_document: default
---
```{r Setup, echo = FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
})

options(dplyr.summarise.inform = FALSE)
```

# EP05 - Precision
Evaluation of Precision of Quantitative Measurement Procedures

This work is based on EP05-A3, October 2014

Simple, elegant code with minimal duplication that can be adapted for other uses.

## Appendix A
Single-Site Study


### Table A1
Glucose Precision Evaluation Measurements (mg/dL)
```{r Table A1}

# Table A1. Glucose Precision Evaluation Measurements (mg/dL)
table_a1 <- tibble(
  day = 1:20,
  run1.rep1 = c(242, 243, 247, 249, 246, 244, 241, 245, 243, 244, 252, 249, 242, 246, 247, 240, 241, 244, 241, 247),
  run1.rep2 = c(246, 242, 239, 241, 242, 245, 246, 245, 239, 246, 251, 248, 240, 249, 248, 238, 244, 244, 239, 240),
  run2.rep1 = c(245, 238, 241, 250, 243, 251, 245, 243, 244, 247, 247, 251, 251, 248, 245, 239, 245, 237, 247, 245),
  run2.rep2 = c(246, 238, 240, 245, 240, 247, 247, 245, 245, 239, 241, 246, 245, 240, 246, 242, 248, 242, 245, 242))

```

#### Normalization
```{r Table A1 Normalization}

table_a1 <- table_a1 %>%
  pivot_longer(cols = starts_with("run"), names_to = "run_rep", values_to = "result") %>%
  separate(run_rep, sep = "\\.", into = c("run", "rep")) %>%
  mutate_at(vars("run", "rep"), ~ str_remove(., "(run)|(rep)")) %>%
  mutate_at(vars(-result), factor)

glimpse(table_a1)

```

### Figure A1
Glucose Precision Evaluation Results
```{r Figure A1}

table_a1 %>%
  ggplot(aes(x = day, y = result)) +
  geom_point(aes(color = run, shape = run)) +
  scale_x_discrete(name = "Test Day", breaks = seq(0, 20, by = 4)) +
  scale_y_continuous(name = "Glucose, mg/dL", breaks = seq(236, 254, by = 2)) +
  scale_color_discrete(name = "", labels = c("Morning Run", "Afternoon Run")) +
  scale_shape_manual(name = "", values = c("square", "circle"), labels = c("Morning Run", "Afternoon Run"))

```


### Table A2
ANOVA Summary Table for the Glucose Example

```{r Table A2}
table_a2 <- table_a1 %>%
  aov(result ~ day + day:run, data = .) %>%
  broom::tidy() %>%
  select(source = term, SS = sumsq, DF = df, MS = meansq) %>%
  column_to_rownames(var = "source")

table_a2
```

### Equations A1-A9

```{r Equations A1-A9}
x_bar <- mean(table_a1$result)

MS_error <- table_a2["Residuals", "MS"]
MS_run <- table_a2["day:run", "MS"]
MS_day <- table_a2["day", "MS"]

DF_error <- table_a2["Residuals", "DF"]
DF_run <- table_a2["day:run", "DF"]
DF_day <- table_a2["day", "DF"]

N <- nrow(table_a1)
n_rep <- nlevels(table_a1$rep)
n_run <- nlevels(table_a1$run)
n_day <- nlevels(table_a1$day)

# If any of these terms are less than zero, carry them forward as zero
V_error <- MS_error  # A1
V_run <- max((MS_run - MS_error)/n_rep, 0)  # A2
V_day <- max((MS_day - MS_run)/(n_run*n_rep), 0)  # A3

S_R <- sqrt(V_error)  # A4
S_WL <- sqrt(V_day + V_run + V_error)  # A5

# Only valid for a 20 x 2 x 2 study
alpha_error <- 0.5
alpha_run <- 0.25
alpha_day <- 0.25

df_R <- N - n_day*n_run  # A6

# A7
df_WL <- round(  
  ((alpha_day * MS_day) + (alpha_run * MS_run) + (alpha_error * MS_error))^2 
  / 
  ((alpha_day * MS_day)^2/DF_day + (alpha_run * MS_run)^2/DF_run + (alpha_error * MS_error)^2/DF_error)
)

CV_R <- S_R / x_bar * 100  # A8
CV_WL <- S_WL / x_bar * 100  # A9

tibble(
  `Sample Description` = "Patient Pool",
  `Mean (mg/dL)` = x_bar,
  `Repeatability` = sprintf("%.2f (%.1f%%)", S_R, CV_R),
  `Within-Laboratory Precision` = sprintf("%.2f (%.1f%%)", S_WL, CV_WL)
) 
```

### Table A3
Inputs for calculating 95% CIs for the Glucose Example

```{r Table A3}
alpha <- 0.05

table_a3 <- tibble(
  value = c("Repeatability", "Within-Laboratory Reproducibility"),
  S = c(S_R, S_WL),
  DF = c(df_R, df_WL),
  chisq_lower = qchisq(1-alpha/2, DF),
  chisq_upper = qchisq(alpha/2, DF))

table_a3 
```

### Equations A10-A11
```{r Equations A10-A11}
table_a3 %>%
  mutate(S_lower = S * sqrt(DF/chisq_lower),  # A10 and A11
         S_upper = S * sqrt(DF/chisq_upper),
         CV_lower = S_lower / x_bar * 100,
         CV_upper = S_upper / x_bar * 100,
         CV_range = sprintf("%.1f%% - %.1f%%", CV_lower, CV_upper)) %>%
  select(value, starts_with("S"), CV_range)

# There is a rounding error in this table, the text reports 0.95% but without rounding, the calculated value is 0.944%
```

## Appendix B
CA19-9 Precision Evaluation Measurements (kU/L). Six samples, P1 to Q6, where P signifies patient sample pool and "Q" signifies QC material.

### B1 - Overview
#### Table B1
```{r Table B1}

# I'm not mad about having to type all these dumb numbers in. Not mad at all
table_b1 <- tibble(
  day = rep(rep(1:5, each = 5), times = 2*3),
  sample = rep(c("P1", "P2", "Q3", "Q4", "P5", "Q6"), each = 25),
  `Site 1` = c(
    12.5,  11.8,  11.6,  11.6,  11.7,  12.2,  11.8,  11.6,  12.2,  11.5,  12.0,  11.9,  11.4,  
    11.6,  11.9,  11.9,  11.1,  12.4,  11.3,  11.5,  13.3,  10.7,  10.3,  10.7,  11.9,   # P1
    45.0,  45.1,  42.4,  42.4,  42.5,  42.4,  41.7,  43.0,  41.8,  41.6,  45.1,  42.7,  41.6,  
    41.0,  42.2,  41.8,  44.4,  42.8,  41.4,  41.8,  41.2,  40.9,  41.1,  40.2,  40.9,   # P2
    59.0,  57.7,  58.1,  56.3,  55.6,  56.5,  56.3,  58.4,  57.3,  56.5,  57.7,  56.7,  58.6,  
    56.1,  55.2,  56.5,  58.4,  55.4,  55.7,  57.7,  56.9,  56.3,  56.3,  55.7,  55.0,   # Q3
    175.0, 167.3, 169.7, 167.8, 170.5, 167.3, 166.7, 172.0, 170.1, 170.5, 167.4, 168.8, 170.5, 
    176.5, 168.2, 173.3, 171.4, 164.1, 166.9, 167.4, 164.8, 167.1, 167.3, 165.8, 165.8,  # Q4
    386.6, 393.9, 392.8, 383.3, 388.0, 389.0, 394.3, 374.2, 379.8, 368.1, 381.0, 380.1, 375.3, 
    375.0, 375.0, 400.2, 386.2, 364.6, 371.1, 371.3, 385.1, 390.6, 386.2, 377.4, 384.9,  # P5
    433.9, 434.6, 424.9, 425.2, 424.9, 420.6, 437.5, 413.6, 416.2, 410.7, 423.0, 412.7, 426.7, 
    435.3, 417.9, 435.9, 424.9, 406.6, 425.8, 427.2, 418.0, 415.8, 409.2, 411.1, 423.2), # Q6
  `Site 2` = c(
    12.8,  12.8,  12.7,  13.2,  11.1,  11.4,  10.7,  11.0,  13.8,  11.4,  14.7,  13.7,  13.2, 
    13.2,  13.1,  13.6,  13.5,  12.8,  12.3,  12.0,  14.1,  12.8,  12.2,  15.8,  13.3,   # P1
    40.6,  39.3,  41.4,  40.3,  37.3,  38.2,  37.5,  38.7,  43.0,  40.7,  42.1,  41.3,  40.5,  
    38.8,  39.9,  40.9,  40.7,  39.6,  40.1,  39.5,  40.8,  40.6,  41.4,  41.4,  37.3,   # P2
    53.9,  55.1,  54.9,  53.7,  51.3,  52.7,  54.4,  49.8,  51.4,  49.0,  55.1,  55.0,  54.1,  
    52.5,  54.5,  55.2,  56.1,  55.2,  55.8,  54.9,  54.5,  55.9,  51.7,  50.0,  54.2,   # Q3
    156.6, 157.7, 155.8, 160.8, 153.6, 162.7, 157.1, 156.1, 149.5, 157.0, 159.9, 163.8, 160.9, 
    164.4, 156.3, 162.2, 164.6, 161.5, 159.2, 157.9, 158.2, 162.9, 162.4, 159.6, 160.0,  # Q4
    380.0, 382.5, 368.5, 379.7, 361.7, 366.8, 371.7, 363.3, 379.1, 366.7, 375.1, 382.1, 371.3, 
    374.3, 370.1, 372.7, 376.1, 363.9, 371.4, 370.8, 365.7, 371.2, 387.5, 385.5, 367.3,  # P5
    410.9, 416.3, 391.8, 390.7, 388.4, 401.2, 409.0, 397.0, 395.7, 384.1, 401.4, 411.9, 392.2, 
    409.3, 391.9, 395.9, 415.7, 400.5, 409.4, 390.8, 410.5, 399.7, 387.9, 398.9, 382.5), # Q6
  `Site 3` = c(
    12.2,  11.7,  11.6,  11.9,  11.6,  12.0,  11.7,  11.9,  12.1,  11.7,  12.0,  11.7,  11.5,  
    11.8,  11.7,  11.5,  11.8,  11.9,  12.3,  12.6,  12.0,  10.5,  10.5,  11.2,  11.1,   # P1
    45.1,  43.8,  42.4,  42.4,  42.1,  42.1,  42.4,  42.4,  43.5,  41.9,  43.9,  42.2,  41.3,  
    41.4,  42.0,  43.1,  43.6,  42.1,  41.3,  42.8,  41.1,  41.0,  40.7,  41.4,  43.9,   # P2
    58.4,  57.9,  57.2,  56.4,  56.1,  56.4,  57.4,  57.9,  57.5,  55.9,  57.2,  57.7,  57.4,  
    56.3,  56.5,  57.5,  56.9,  55.6,  56.3,  56.1,  56.6,  56.3,  56.0,  55.7,  57.0,   # Q3
    171.2, 168.5, 168.8, 167.6, 170.5, 167.0, 169.4, 171.1, 168.8, 169.4, 168.1, 169.7, 173.5,
    174.9, 167.8, 172.4, 167.8, 165.5, 165.9, 166.0, 166.0, 167.2, 166.6, 168.2, 168.8,  # Q4
    390.3, 393.4, 388.1, 386.2, 378.1, 391.7, 384.3, 377.0, 380.4, 371.6, 380.6, 377.7, 375.2, 
    387.6, 373.2, 393.2, 375.4, 367.9, 378.1, 378.3, 387.9, 388.4, 381.8, 382.7, 383.7,  # P5
    434.3, 429.8, 425.1, 422.9, 417.8, 429.1, 425.6, 414.9, 419.6, 414.3, 417.9, 419.7, 431.0, 
    435.6, 422.6, 430.4, 415.8, 416.2, 421.9, 415.2, 416.9, 412.5, 410.2, 418.0, 415.2), # Q6
)

```

##### Normalization

```{r Table B1 Normalization}
table_b1 <- table_b1 %>%
  pivot_longer(cols = starts_with("Site"), names_to = "site", values_to = "result") %>%
  mutate(sample = factor(sample, levels = c("P1", "P2", "Q3", "Q4", "P5", "Q6"))) %>%
  mutate_at(vars(-result), as.factor)
```

### B2 - Screen for Anomolies
#### Figure B1
Screen for Anomalies - All Samples. The vertical axis represents a robust Levey-Jennings scale.

```{r Figure B1}

table_b1 %>%
  group_by(sample, site) %>%
  mutate(deviation = (result - median(result)) / mad(result)) %>%  # This doesn't match up?
  ggplot(aes(x = sample, y = deviation)) +
  geom_jitter(height = 0) +
  geom_hline(yintercept = c(-4, -3, 0, 3, 4), lty = 2, lwd = 0.25) +
  facet_grid(~ site) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "Deviation from Median in MAD Units", breaks = seq(-4, 4, by = 2))

```

#### Figure B2
```{r Figure B2}

table_b1 %>%
  filter(sample == "P1") %>%
  group_by(site) %>%
  mutate(meas_n = 1:n()) %>%
  ggplot(aes(x = meas_n, y = result)) +
  geom_jitter(height = 0) +
  facet_wrap(~ site) +
  scale_x_continuous(name = "", breaks = seq(3, 24, by = 5), labels = paste0("D", 1:5)) +
  scale_y_continuous(name = "CA19-9, kU/L - Sample P1", breaks = seq(9, 17, 2)) +
  
  # Add mean and Grubbs' line
  stat_summary(geom = "hline", fun = mean, aes(x = 1, yintercept = ..y..)) +
  stat_summary(geom = "hline", fun = ~ mean(.) + 3.135 * sd(.), aes(x = 1, yintercept = ..y..)) +
  stat_summary(geom = "hline", fun = ~ mean(.) - 3.135 * sd(.), aes(x = 1, yintercept = ..y..)) +
  
  # Add median line and +/- 3, 4 MAD
  stat_summary(geom = "hline", fun = median, aes(x = 1, yintercept = ..y..), lty = 2) +
  stat_summary(geom = "hline", fun = ~ median(.) + 3 * mad(.), aes(x = 1, yintercept = ..y..), lty = 2, lwd = 0.25) +
  stat_summary(geom = "hline", fun = ~ median(.) - 3 * mad(.), aes(x = 1, yintercept = ..y..), lty = 2, lwd = 0.25) +
  stat_summary(geom = "hline", fun = ~ median(.) + 4 * mad(.), aes(x = 1, yintercept = ..y..), lty = 2, lwd = 0.25) +
  stat_summary(geom = "hline", fun = ~ median(.) - 4 * mad(.), aes(x = 1, yintercept = ..y..), lty = 2, lwd = 0.25)

```
### B3 - Analysis of Variance
#### Table B2

```{r Table B2}

variance_across_sites <- function(sample) {
  
  # Function to perform the variance component calculations, like Table B2 and the top section of Table B3
  # sample: the name of the test sample, stored in a column called 'sample'
  
  # returns: a list with the following values:
  #   sample: name of the sample (as passed in the function call)
  #   mean: mean value of that sample
  #   N: number of measurements in the dataset
  #   df_R: degrees of freedom associated with quantifying the repeatability
  #   V_error: variance associated with repeatability of the assay
  #   %_error: V_error as a percentage of total variance
  #   SD_error: standard deviation associated with repeatability of the assay
  #   CV_error: standard deviation expressed as a percentage of mean
  #   df_WL: degrees of freedom of within-laboratory precision estimate, calculated via Satterthwaite approximation
  #   V_day, %_day, SD_day, CV_day: as above, except WRT the day-to-day variation of the assay
  #   df_REP: degrees of freedom of reproducibility precision estimate, calculated via Satterthwaite approximation
  #   V_site, %_site, SD_site, CV_site: as above, except WRT the site-to-site variation of the assay

  df <- filter(table_b1, sample == (!! sample))  # We have to be explicit here so we compare the function parameter 'sample' to the column 'sample'
  
  aov_table <- df %>%
    aov(result ~ site + site:day, data = .) %>%
    broom::tidy() %>%
    select(source = term, SS = sumsq, DF = df, MS = meansq) %>%
    column_to_rownames(var = "source")
  
  x_bar <- mean(df$result)
  
  MS_error <- aov_table["Residuals", "MS"]
  MS_day <- aov_table["site:day", "MS"]
  MS_site <- aov_table["site", "MS"]
  
  DF_error <- aov_table["Residuals", "DF"]
  DF_day <- aov_table["site:day", "DF"]
  DF_site <- aov_table["site", "DF"]
  
  N <- nrow(df)
  n_day <- nlevels(df$day)
  n_site <- nlevels(df$site)
  n_rep <- N / (n_day * n_site)
  
  # If any of these terms are less than zero, carry them forward as zero
  V_error <- MS_error 
  V_day <- max((MS_day - MS_error)/n_rep, 0) 
  V_site <- max((MS_site - MS_day)/(n_day*n_rep), 0) 
  
  S_R <- sqrt(V_error) 
  S_WL <- sqrt(V_day + V_error) 
  S_REP <- sqrt(V_site + V_day + V_error)
  
  # Only valid for a 3 x 5 x 5 study
  alpha_day <- 0.5
  alpha_error <- 0.5
  beta_site <- 0.25
  beta_day <- 0.25
  beta_error <- 0.5
  
  df_R <- N - n_site*n_day
  
  df_WL <- round(  
    (alpha_day * MS_day + alpha_error * MS_error)^2 
    / 
    ((alpha_day * MS_day)^2/DF_day + (alpha_error * MS_error)^2/DF_error)
  )
  
  df_REP <- round(  
    ((beta_site * MS_site) + (beta_day * MS_day) + (beta_error * MS_error))^2 
    / 
    ((beta_site * MS_site)^2/DF_site + (beta_day * MS_day)^2/DF_day + (beta_error * MS_error)^2/DF_error)
  )
  
  CV_R <- S_R / x_bar * 100
  CV_WL <- S_WL / x_bar * 100
  CV_REP <- S_REP / x_bar * 100
  
  list(
    sample = sample,
    mean = x_bar,
    N = nrow(df),
    df_R = df_R,
    V_error = V_error,
    `%_error` = V_error / (V_error + V_day + V_site) * 100,
    SD_R = S_R,
    CV_R = S_R/x_bar * 100,
    df_WL = df_WL,
    V_day = V_day,
    `%_day` = V_day / (V_error + V_day + V_site) * 100,
    SD_WL = S_WL,
    CV_WL = S_WL/x_bar * 100,
    df_REP = df_REP,
    V_site = V_site,
    `%_site` = V_site / (V_error + V_day + V_site) * 100,
    SD_REP = S_REP,
    CV_REP = S_REP/x_bar * 100
  )
}
  
samples <- levels(table_b1$sample)
  
table_b2 <- map_dfr(samples, ~ variance_across_sites(.))

# The values for P5 don't match up with the text, but I can't find a mistake in the data
table_b2 %>%
  select(sample, mean, V_error, `%_error`, V_day, `%_day`, V_site, `%_site`)

```

#### Table B3
##### Combined Dataset
```{r Table B3, Combined Dataset}
table_b2 %>%
  select(sample, mean, SD_R, CV_R, SD_WL, CV_WL, SD_REP, CV_REP)
```

##### By Site
```{r Single Site Variance Function}

variance_at_single_site <- function(site, sample) {

  # Function to perform the variance component calculations, like Table B3, at a single site
  # sample: sample identifier, stored in a column called 'sample'
  # site: site identifier, stored in a column called 'site'
  
  # returns: a list with the following values:
  #   sample: name of the sample (as passed in the function call)
  #   mean: mean value of that sample
  #   N: number of measurements in the dataset
  #   df_R: degrees of freedom associated with quantifying the repeatability
  #   V_error: variance associated with repeatability of the assay
  #   %_error: V_error as a percentage of total variance
  #   SD_error: standard deviation associated with repeatability of the assay
  #   CV_error: standard deviation expressed as a percentage of mean
  #   df_WL: degrees of freedom of within-laboratory precision estimate, calculated via Satterthwaite approximation
  #   V_day, %_day, SD_day, CV_day: as above, except WRT the day-to-day variation of the assay

  df <- filter(table_b1, 
               sample == (!! sample),
               site == (!! site))
  
  aov_table <- df %>%
    aov(result ~ day, data = .) %>%
    broom::tidy() %>%
    select(source = term, SS = sumsq, DF = df, MS = meansq) %>%
    column_to_rownames(var = "source")
  
  x_bar <- mean(df$result)
  
  MS_error <- aov_table["Residuals", "MS"]
  MS_day <- aov_table["day", "MS"]

  DF_error <- aov_table["Residuals", "DF"]
  DF_day <- aov_table["day", "DF"]

  N <- nrow(df)
  n_day <- nlevels(df$day)
  n_rep <- N / n_day
  
  # If any of these terms are less than zero, carry them forward as zero
  V_error <- MS_error 
  V_day <- max((MS_day - MS_error)/n_rep, 0) 
  
  S_R <- sqrt(V_error) 
  S_WL <- sqrt(V_day + V_error)
  
  # Only valid for a 5 x 5 study
  alpha_day <- 0.5
  alpha_error <- 0.5

  df_R <- N - n_day
  
  df_WL <- round(  
    ((alpha_day * MS_day) + (alpha_error * MS_error))^2 
    / 
    ((alpha_day * MS_day)^2/DF_day + (alpha_error * MS_error)^2/DF_error)
  )
  
  CV_R <- S_R / x_bar * 100
  CV_WL <- S_WL / x_bar * 100
  
  list(
    site = site,
    sample = sample,
    mean = x_bar,
    N = nrow(df),
    df_R = df_R,
    V_error = V_error,
    `%_error` = V_error / (V_error + V_day) * 100,
    SD_R = S_R,
    CV_R = S_R/x_bar * 100,
    df_WL = df_WL,
    V_day = V_day,
    `%_day` = V_day / (V_error + V_day) * 100,
    SD_WL = S_WL,
    CV_WL = S_WL/x_bar * 100
  )
}

sample_site <- expand.grid(
  unique(table_b1$sample),
  unique(table_b1$site)
)

table_b3 <- pmap_dfr(sample_site, ~ variance_at_single_site(sample = ..1, site = ..2)) %>%
  select(site, sample, mean, SD_R, CV_R, SD_WL, CV_WL) 

```

###### Site 1
```{r Table B3, Site 1}

filter(table_b3, site == "Site 1")

```

###### Site 2
```{r Table B3, Site 2}

filter(table_b3, site == "Site 2")

```

###### Site 3
```{r Table B3, Site 3}

filter(table_b3, site == "Site 3")

```

#### Table B4
This doesn't match with the text!
```{r Table B4}

alpha <- 0.05

means <- select(table_b2, sample, mean)

# These numbers don't match up with the text!
# There is likely a mistake in here somewhere, but I can't find it
table_b4 <- table_b2 %>%
  mutate(SD_R_lower = SD_R * sqrt(df_R / qchisq(1-alpha/2, df_R)),
         SD_R_upper = SD_R * sqrt(df_R / qchisq(alpha/2, df_R)),
         SD_R_r = sprintf("%.3f - %.3f", SD_R_lower, SD_R_upper),
         
         CV_R_lower = SD_R_lower / mean * 100,
         CV_R_upper = SD_R_upper / mean * 100,
         CV_R_r = sprintf("%.1f%% - %.1f%%", CV_R_lower, CV_R_upper),
         
         SD_WL_lower = SD_WL * sqrt(df_WL / qchisq(1-alpha/2, df_WL)),
         SD_WL_upper = SD_WL * sqrt(df_WL / qchisq(alpha/2, df_WL)),
         SD_WL_r = sprintf("%.3f - %.3f", SD_WL_lower, SD_WL_upper),
         
         CV_WL_lower = SD_WL_lower / mean * 100,
         CV_WL_upper = SD_WL_upper / mean * 100,
         CV_WL_r = sprintf("%.1f%% - %.1f%%", CV_WL_lower, CV_WL_upper),
         
         SD_REP_lower = SD_REP * sqrt(df_REP / qchisq(1-alpha/2, df_REP)),
         SD_REP_upper = SD_REP * sqrt(df_REP / qchisq(alpha/2, df_REP)),
         SD_REP_r = sprintf("%.3f - %.3f", SD_REP_lower, SD_REP_upper),
         
         CV_REP_lower = SD_REP_lower / mean * 100,
         CV_REP_upper = SD_REP_upper / mean * 100,
         CV_REP_r = sprintf("%.1f%% - %.1f%%", CV_REP_lower, CV_REP_upper)) 

table_b4 %>%
  select(Sample = sample, 
         `Repeatability SD` = SD_R_r, 
         `Repeatability %CV` = CV_R_r, 
         `Within-Lab Precision SD` = SD_WL_r, 
         `Within Lab Precision %CV` = CV_WL_r, 
         `Reproducibility SD` = SD_REP_r, 
         `Reproducibility %CV` = CV_REP_r)

```

#### Figure B3

```{r Figure B3}

table_b4_long <- table_b4 %>%
  select(sample, mean, 
         starts_with("SD_"), starts_with("CV_"), 
         -ends_with("_upper"), -ends_with("_lower"), 
         -ends_with("_r", ignore.case = FALSE)) %>%
  pivot_longer(cols = c(starts_with("CV_"), starts_with("SD_")), names_to = "type", values_to = "value")

# Because variance is fit to a relatively uncommon function, we have to fit it 
# by brute force, using nonlinear least squares nls()

models <- table_b4_long %>%
  mutate(Var = value^2) %>%
  filter(str_detect(type, "SD_")) %>%
  group_by(type) %>% 
  do(model = list(
    nls(data = .,
        formula = log(Var) ~ C * log(A + B * mean),  # We rearrange this to make it play nice with nls()
        start = c("A" = 0, "B" = 0.02, "C" = 1))  # Initial guesses for the parameters
  )) %>%
  mutate(coefficients = map(model, ~ coef(.))) %>%
  unnest_wider(coefficients)

table_b4_long %>%
  filter(str_detect(type, "CV_")) %>%
  mutate(type = type %>%
           fct_recode(Repeatability = "CV_R", `Within-Laboratory Precision` = "CV_WL", Reproducibility = "CV_REP") %>%
           fct_relevel("Repeatability", "Within-Laboratory Precision", "Reproducibility")) %>%
  ggplot(aes(x = mean, y = value)) +
  geom_point(aes(color = type, shape = type), size = 3) +
  # Adding these functions in manually -- there may be a better way
  # Function returns Var, CV = 100 * sqrt(Var)/mean
  stat_function(fun = ~ 100*sqrt((models$A[[1]] + models$B[[1]] * .x)^models$C[[1]])/.x, color = "#F8766D") +
  stat_function(fun = ~ 100*sqrt((models$A[[2]] + models$B[[2]] * .x)^models$C[[2]])/.x, color = "#619CFF") +
  stat_function(fun = ~ 100*sqrt((models$A[[3]] + models$B[[3]] * .x)^models$C[[3]])/.x, color = "#00BA38") +
  scale_x_continuous(breaks = c(12, 50, 150, 400), minor_breaks = c(), name = "CA19-9, kU/L (Log Scale)") +
  scale_y_continuous(breaks = seq(0, 10, by = 2), name = "%CV") +
  scale_color_discrete(name = "") +
  scale_shape_manual(values = c("triangle", "circle", "square"), name = "") +
  coord_trans(x = "log10")

```