--- 
title: "Method Evaluation for Tidyverse Users"
author: "Colin Douglas"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "Reproducing (in tidyverse-style R) the worked examples from the CLSI's guidelines on evaluation of clinical laboratory measurement procedures."
---

# Prerequisites

Placeholder



<!--chapter:end:index.Rmd-->

# Introduction {#intro}

This guide intends to reproduceworked examples from the appendices of CLSI guidelines in clean, easy-to-read tidyverse-style R. It makes no assurances that the analysis is appropriate or extendible, just that the results are reproducible.

<!--chapter:end:01-intro.Rmd-->

# EP17 - Detection Capabilities
Evaluation of Detection Capability for Clinical Laboratory Measurement Procedures

This work is based on EP17-A2, June 2012.

## Appendix A
Evaluation of Limits of Blank and Detection by the Classical Approach

```{r Setup, echo = FALSE}
library(tidyverse)
```


```{r Table A1}
table_a1 <- tibble(
  day = rep(1:3, each = 4),
  replicate = rep(1:4, times = 3)
)

<!--chapter:end:02-EP17.Rmd-->

# Literature

Here is a review of existing methods.

<!--chapter:end:02-literature.Rmd-->

# Methods

We describe our methods in this chapter.

<!--chapter:end:03-method.Rmd-->

# Applications

Some _significant_ applications are demonstrated in this chapter.

## Example one

## Example two

<!--chapter:end:04-application.Rmd-->

# Final Words

We have finished a nice book.

<!--chapter:end:05-summary.Rmd-->

`r if (knitr::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

